Bootstrap Confidence Intervals for Sharp Regression Discontinuity Designs with the Uniform Kernel

### 2016-03-01 main_paper.tex
Make some small edits to the application of the main paper

\begin{abstract}\noindent
  This paper develops a novel bootstrap procedure to obtain robust
  bias-corrected confidence intervals in regression discontinuity designs that
  use the uniform kernel. The procedure uses a residual bootstrap to estimate
  the bias of the RD estimator, which is then removed from the original
  estimator.  The bias-corrected estimator is then bootstrapped itself to
  generate valid confidence intervals. This procedure is valid for empirically
  relevant bandwidths when the treatment effect is estimated with a local linear
  model and the residual bootstraps use a second order polynomial; validity
  holds under similar conditions to Calonico, Cattaneo and Titiunik's (2014,
  \textit{Econometrica}) analytical correction.
\end{abstract}

@@

The paper is organized as follows. Section~\ref{background} describes the basic
RD approach, its usual implementation, and the explicit analytical bias
correction approach in the literature. Section~\ref{boot} presents our proposed
bootstrap bias corrected RD algorithm and discusses its asymptotic
properties. Simulation evidence that the bootstrap procedure provides valid CIs
and its relative performance to the analytical bias correction are presented in
Section~\ref{sim}.
Finally, Section~\ref{conclusion} concludes.

@@

This section presents evidence from Monte Carlo simulations that the bootstrap
procedures proposed in Section~\ref{boot} produce valid, robust confidence
intervals similar to the ones obtained by the analytical procedures established
in CCT. The bootstrap CIs obtained compare favorably to the analytical
alternative, with coverage slightly closer to nominal coverage and shorter
length of the intervals in the specifications implemented.\footnote{%
  All of the simulations were carried out in the R programming language
  \citep{R} and rely on the \textit{rdrobust} \citep{rdrobust}, \textit{foreach}
  \citep{foreach}, \textit{doParallel} \citep{doparallel}, and \textit{doRNG}
  \citep{dorng} packages.} %

@@

\section{Conclusion}\label{conclusion}
This paper proposes a novel bootstrap procedure to obtain robust bias-corrected confidence intervals in regression discontinuity designs.
The approach proposed builds upon the developments and intuition advanced by CCT and is based on a first-order bias correction.
 We exploit CCT's theoretical insight through a new residual bootstrap. In particular, we propose estimating the local linear model as usual,
then estimating a local second order polynomial and generating bootstrap datasets by resampling the residuals of that polynomial.

The bootstrap sample is used to estimate the bias of the RD estimator, which is then removed from the original estimator.
 This approach is described in detail by our Algorithm~\ref{Alg1} and the resulting bias corrected estimator is shown to be asymptotically normal with mean zero in our
Theorem~\ref{T1} under AMSE-optimal bandwidth rates.

Just as in CCT, our bias correction step introduces additional variability.  So we propose then bootstrap the bias-corrected estimator itself to
 generate valid confidence intervals. This procedure, which requires bootstrapping the datasets produced by an initial bootstrap procedure, is described in Algorithm~\ref{Alg2},
 and the resulting confidence intervals are shown to be asymptotically valid in Theorem~\ref{T2}.

 Hence, our procedure provides an alternative to the plug-in analytical methods in the literature and is simple to implement generating robust confidence intervals.

Simulation evidence is presented that the proposed bootstrap bias correction and confidence intervals have similar empirical coverage and average interval length relative to
 the analytical alternatives proposed by CCT, making it a suitable option to practitioners.

>>>>

\begin{abstract}\noindent
  This paper develops a novel bootstrap procedure to obtain robust
  bias-corrected confidence intervals in regression discontinuity designs that
  use the uniform kernel. The procedure uses a residual bootstrap to estimate
  the bias of the RD estimator, which is then removed from the original
  estimator.  The bias-corrected estimator is then bootstrapped itself to
  generate valid confidence intervals. This procedure is valid for empirically
  relevant bandwidths when the treatment effect is estimated with a local linear
  model and the residual bootstraps use a second order polynomial; validity
  holds under similar conditions to Calonico, Cattaneo and Titiunik's (2014,
  \textit{Econometrica}) analytical correction. This paper also provides
  simulation evidence that our method is as accurate as the analytical
  corrections and we demonstrate its use through a reanalysis of Ludwig and
  Miller's (2008) Head Start dataset.
\end{abstract}

@@

The paper is organized as follows. Section~\ref{background} describes the basic
RD approach, its usual implementation, and the explicit analytical bias
correction approach in the literature. Section~\ref{boot} presents our proposed
bootstrap bias corrected RD algorithm and discusses its asymptotic
properties. Simulation evidence that the bootstrap procedure provides valid CIs
and its relative performance to the analytical bias correction are presented in
Section~\ref{sim} and Section~\ref{application} demonstrates the estimator's
usage by applying it to the Head Start dataset used
by~\cite{ludwig2007}.\footnote{%
  The simulations and empirical analysis were carried out in the R programming
  language \citep{R} and rely on the \textit{rdrobust} \citep{rdrobust},
  \textit{doParallel} \citep{doparallel}, \textit{foreach} \citep{foreach}, and
  \textit{doRNG} \citep{dorng} packages.} %
Finally, Section~\ref{conclusion} concludes.

@@

This section presents evidence from Monte Carlo simulations that the bootstrap
procedures proposed in Section~\ref{boot} produce valid, robust confidence
intervals similar to the ones obtained by the analytical procedures established
in CCT. The bootstrap CIs obtained compare favorably to the analytical
alternative, with coverage slightly closer to nominal coverage and shorter
length of the intervals in the specifications implemented.

@@

\section{Application}\label{application}

In this section, we apply the bootstrap procedure to the data used in
\cite{ludwig2007}.\footnote{%
  The data is publicly available from
  \url{http://faculty.econ.ucdavis.edu/faculty/dlmiller/statafiles}.} %
In their paper, the effects of Head Start application assistance on health and
schooling were investigated under a sharp RD design.

In 1965, the Head Start program was established to help poor children age three
to five and their families. The program elements include parent involvement,
nutrition, social services, mental health services and health services. To
promote this program in the most needing area, the Office of Economic
Opportunity provided grant-writing assistance to the poorest 300 counties in the
United States based on the 1960 poverty rate. So the poverty rate of the 300th
poorest county serves as a sharp cutoff of treatment. It is shown in
\cite{ludwig2007} that the 228 ``treatment'' counties with poverty rates 10
percentage points above this cutoff have average Head Start spending per
four-year-old as twice of that for 349 ``control'' counties with poverty rates
10 percentage points below this cutoff.

\cite{ludwig2007} utilize this fact to estimate the ``intent-to-treat'' effect:
the effect of proposal developing assistance on health and schooling. They use
mortality as their health measure and use data from the National Vital
Statistics System of the National Center for Health Statistics, which provide
information on cause of death and age at death. \cite{ludwig2007} limited the
causes of death to those which could be affected by Head Start health services
and found a large drop in mortality rates of children five to nine years of age
over the period of 1973--1983. They also found some evidence for a positive
effect on schooling from decennial census data.

We reestimate the ATE on health and schooling with robust procedures. To be
specific, we apply both the robust procedure proposed by CCT and the bootstrap
procedure introduced in this paper.  Our bootstrap estimator uses the
AMSE-optimal bandwidths proposed by CCT and the uniform kernel. In the bootstrap
procedure, we use 500 bootstraps for bias correction and 999 to calculate the
confidence intervals.\footnote{%
  These are the same design choices made for the Monte Carlo study.} %
The analytical estimator using CCT's bias correction and variance estimator use
their AMSE-optimal bandwidths as well.

\begin{table}[t]
  \centering
  \begin{tabular}{lrr@{, }rr@{}rlr}
    \toprule
                      & ATE     & \multicolumn{2}{r}{95\% CI}           & $h$&      & $b$   & p-value \\
    \midrule
    LM (2007)         & --1.895 & (--3.930                    & 0.139)   &  9&     &       & 0.036   \\
    LM (2007)         & --1.198 & (--2.561                    & 0.165)   & 18&     &       & 0.081   \\
    LM (2007)         & --1.114 & (--2.138                    & --0.090) & 36&     &       & 0.027   \\
    CCT               & --3.795 & (--7.037                    & --0.554) &  3&.888 & 6.807 &         \\
    Resid.\ bootstrap & --3.792 & (--6.512                    & --0.262) &  3&.888 & 6.807 &         \\
    \bottomrule
  \end{tabular}
  \caption{The effect of Head Start assistance on mortality. The first three
    rows come from Table 3 in \cite{ludwig2007} except ``95\% CI,'' which is
    calculated using the conventional asymptotic interval estimator.
    The last two rows list results from two robust procedures.}
  \label{tbl:2}
\end{table}

For completeness, we also report the
original results in \cite{ludwig2007}.\footnote{%
  One minor issue arose in reproducing Ludwig and Miller's results: their paper
  presents results for the \emph{triangular} kernel, but we were only able to
  recover their ATE estimate using the \emph{uniform} kernel. Results for the
  triangular and uniform kernel were essentially the same, and this discrepency
  does not affect Ludwig and Miller's conclusions.} %
Since our research focuses on confidence intervals, we calculate and report
the conventional unadjusted RD confidence interval for each of the bandwidths
used by \cite{ludwig2007}. (Ludwig and Miller use a range of bandwidths
for their analysis.) They also use a paired bootstrap algorithm to generate
p-values, and we report those p-values for completeness.

Table \ref{tbl:2} shows the results of the Head Start program on mortality of
children five to nine years of age.\footnote{%
  We, like Ludwig and Miller, focus on the 1973--1983 period.} %
Instead of choosing an optimal bandwidth, \cite{ludwig2007} adopted three
candidate bandwidths 9, 18, and 36. Their estimates indicate that Head Start
assistance lowers the targeted mortality rate by --1.895, --1.198 and --1.114
respectively, which are not very sensitive to the choice of bandwidths in this
range. These ATEs are also significantly different from zero (with p-value
0.036, 0.081 and 0.027) based on Ludwig and Miller's percentile-$T$ bootstrapped
p-value. The statistical inference changes when conventional analytical
confidence interval is used, which includes zero for bandwidth 9 and 18.

Results from the two robust procedures are similar to each other but greatly
differ from the original estimates. The estimated ATEs are as high as -3.795
from CCT and -3.792 from the bootstrap. Both are significantly different from
zero, though the confidence intervals are also very wide, which is likely to be
due to the much smaller bandwidths used relative to the other estimators
($h=3.888, b=6.807$).

Table \ref{tbl:3} presents the effect of the program on schooling for the cohort
aged 18-24 in 1990. The measurement of schooling is the fraction of people with
high school or more in Panel A and the fraction of people with some college or
more in Panel B. A bandwidth of 7 is used in \cite{ludwig2007}. Their estimates
suggest that Head Start assistance increases the fraction of people with high
school or more by 3\% and the fraction of people with some college or more by
3.7\%. Both are significantly different from zero based on both their p-value
and the conventional analytical confidence interval.

The two robust procedures are again very similar to each other and give slightly
larger point estimates in both panels. In Panel A, the ATE increases from 0.030
to 0.055 (CCT) and 0.054 (bootstrap). In Panel B, the ATE increases from 0.037
to 0.051 (CCT) and 0.052 (bootstrap). The confidence intervals tend to shift up
as well.  In contrast to the striking differences between the conventional
estimates bias-corrected procedures for the mortality estimates, the differences
in Table~\ref{tbl:3} are much smaller.

\begin{table}[t]
  \centering
  \begin{tabular}{lrr@{, }rr@{}rlr}
    \toprule
              & ATE   & \multicolumn{2}{r}{95\% CI} & $h$    &   & $b$  & p-value     \\
    \midrule                                                                          \\
    \multicolumn{5}{l}{Fraction ``high school or more'' (Panel A)}                    \\
    \midrule
    LM (2007) & 0.030 & (0.003                      & 0.057) & 7&     &  & 0.032    \\
    CCT       & 0.055 & (0.014                      & 0.096) & 3&.671 & 8.618       \\
    Bootstrap & 0.054 & (0.013                      & 0.096) & 3&.671 & 8.618       \\\\
    \multicolumn{5}{l}{Fraction ``some college or more'' (Panel B)}                   \\
    \midrule
    LM (2007) & 0.037 & (0.002                      & 0.073) & 7&     &  & 0.032 \T \\
    CCT       & 0.051 & (0.004                      & 0.099) & 5&.076 & 10.251      \\
    Bootstrap & 0.052 & (0.001                      & 0.094) & 5&.076 & 10.251      \\
    \bottomrule
  \end{tabular}
  \caption{The effect of Head Start assistance on education for cohort
    18-24 in 1990. Panel A uses the fraction of people with high school
    or more as dependent variable. Panel B uses the fraction of people
    with some college or more as dependent variable. The first row in
    each panel comes from Table 4 in \cite{ludwig2007} except ``95\% CI,''
    which is calculated using the conventional asymptotic interval estimator.
    The last two rows in
    each panel list results from two robust procedures.}
  \label{tbl:3}
\end{table}

To briefly summarize, the bootstrap procedure performs similarly to the robust
estimator proposed by CCT in the above applications. Both provide somewhat
dissimilar answers from the classical point and interval estimators that do
not account for the estimator's bias, but our estimates largely support the
direction and the statistical significance of Ludwig and Miller's (2007)
empirical findings.

\section{Conclusion}\label{conclusion}

This paper proposes a novel bootstrap procedure to obtain robust bias-corrected
confidence intervals in sharp regression discontinuity designs using a uniform
kernel. The approach proposed builds upon the developments and intuition
advanced by CCT and is based on a first-order bias correction. We exploit CCT's
theoretical insight through a new residual bootstrap. In particular, we propose
estimating the local linear model as usual, then estimating a local second order
polynomial and generating bootstrap datasets by resampling the residuals of that
polynomial. This bootstrap allows the bias of the linear model to be estimated
and removed, and the bootstrap can be repeated to accurately estimate the
sampling distribution of the bias-corrected estimator.

Bootstrap procedures are appealing in this setting because they are relatively
easy for applied researchers to modify to account for new and unusual dependence
structures or functional forms. The variance adjustment was proven to be correct
under the assumption of i.i.d.\ data, but could easily be extended to handle
cross-sectional or time series dependence by using the appropriate resampling
strategy on the second order polynomial's residuals. For the analytical methods,
however, new formulas need to be derived to accomodate new features of the
DGP. To fully take advantage of the bootstrap's flexibility, though, these
results need to be extended to other kernels, design strategies (e.g. fuzzy and
kink RD designs) and other more realistic dependence structures. All of these
are the subject of ongoing research.
